---
title: "RV144 Aids trials"
author: "instr"
date: "April 13, 2016"
output: html_document
---


# Thailand Aids study
```{r include=FALSE}
require(tigerstats)
```

## Introduction

In the late 1980s in Thailand, there was a dramatic increase in the prevalence of infection with the human immunodeficiency virus type 1 (HIV-1) in sentinel surveillance cohorts.1-3 Initially, these groups consisted of injection-drug users and commercial sex workers; they were subsequently expanded to include persons in the general population. By 1995, the overall seroprevalence of HIV-1 reached a peak of 3.7% among conscripts in the Royal Thai Army and of 12.5% among conscripts from Northern Thailand.2,4,5 The Thai Ministry of Public Health responded with an effective HIV-prevention campaign, and the number of new HIV-1 infections per year decreased from an estimated 143,000 in 1990 to 14,000 in 2007.2,4,6-9 The persistence of new infection despite these measures led public health officials to conclude that an HIV vaccine, within the context of a broader HIV-prevention program, was needed for better control of the epidemic.



## Methods and data

Study Design and Population
This study was a community-based, randomized, multicenter, double-blind, placebo-controlled efficacy trial of the prime–boost combination of vaccines containing ALVAC-HIV (vCP1521) (Sanofi Pasteur) and AIDSVAX B/E (Global Solutions for Infectious Diseases). For details regarding the vaccines and placebo, see the Supplementary Appendix, available with the full text of this article at NEJM.org. The study was designed to evaluate two coprimary end points: the prevention of HIV-1 infection and the effect of vaccination on the early viral load after infection. The trial was conducted through facilities of the Thai Ministry of Public Health in Rayong and Chon Buri provinces. From September 2003 through December 2005, a total of 16,402 volunteers were enrolled.
Thai men and women who were between the ages of 18 and 30 years and who were not infected with HIV were recruited from the community without regard to HIV risk (i.e., community risk). Written informed consent was obtained from all volunteers, who were required to pass a written test of understanding. Women were counseled to practice effective contraception until 3 months after the last vaccination; pregnant and breast-feeding women were excluded.

###Study Design and Population
This study was a community-based, randomized, multicenter, double-blind, placebo-controlled efficacy trial of the prime–boost combination of vaccines containing ALVAC-HIV (vCP1521) (Sanofi Pasteur) and AIDSVAX B/E (Global Solutions for Infectious Diseases). For details regarding the vaccines and placebo, see the Supplementary Appendix, available with the full text of this article at NEJM.org. The study was designed to evaluate two coprimary end points: the prevention of HIV-1 infection and the effect of vaccination on the early viral load after infection. The trial was conducted through facilities of the Thai Ministry of Public Health in Rayong and Chon Buri provinces. From September 2003 through December 2005, a total of 16,402 volunteers were enrolled.
Thai men and women who were between the ages of 18 and 30 years and who were not infected with HIV were recruited from the community without regard to HIV risk (i.e., community risk). Written informed consent was obtained from all volunteers, who were required to pass a written test of understanding. Women were counseled to practice effective contraception until 3 months after the last vaccination; pregnant and breast-feeding women were excluded.

###Primary End Points
HIV-1 Infection
HIV-1 infection was diagnosed in 132 subjects (56 in the vaccine group and 76 in the placebo group) during 52,985 person-years of follow-up in the intention-to-treat analysis, in 86 subjects (36 in the vaccine group and 50 in the placebo group) during 36,720 person-years of follow-up in the per-protocol analysis, and in 125 subjects (51 in the vaccine group and 74 in the placebo group) during 52,985 person-years of follow-up in the modified intention-to-treat analysis. One subject in the placebo group who was identified by hospital record as being seropositive for HIV after dying from Pneumocystis jirovecii pneumonia was included in the analysis before the unblinding of the study. This diagnosis of HIV-1 infection was the only one that occurred outside planned procedures.
We create the following data set

```{r}
groupsize = 16402
subgp = groupsize/2
vaccine = c(subgp-51,51)
placebo =  c(subgp-74,74)
results = as.table(cbind(vaccine,placebo))
rownames(results) <- c("none","aids")

```


## Results
<!-- You are to add descriptive comments to explain to the reader what the plot is showing !-->

### Graphical results
<!--Display graphical evidence as to the effectiveness of the vaccine with explanations!-->
```{r}
barchartGC(results,type = "percent")
```

### Numerical Results

<!--Display numerical evidence as to the effectiveness of the vaccine with explanations !--> 

```{r}
colPerc(results)
rowPerc(results)
```




## Inferential results

<!--Display inferential evidence as to the effectiveness of the vaccine with explanations and interpretations!-->
<<<<<<< HEAD

```{r}
chisqtestGC(results, graph = TRUE)
```


## Conclusions

<Summarise your results as to the effectiveness of the vaccine with explanations that could be understood by your mom>

After analyzing the data both graphically and numerically, it was determined that more vaccine recipients remained HIV negative than placebo recipients. Directly related to that data, we found that out of those individuals who did end up contracting the virus, a significantly greater amount of them had taken the placebo. This led us to believe that the vaccine might be effective.  

In order to asses that possibility, we used both the Chi-Square test and Fisher's Exact test. Since both tests yielded p-values lower the typically accepted 5% cut-off value, we concluded that there **is** in fact a relationship between the vaccine and the contraction of the virus. More specifically, those who received the vaccine were *less* likely to contract the virus. Thus, the vaccine does appear to be somewhat effective in reducing one's chances of contracting HIV and developing Aids.
=======
 
```{r}
CHI <- chisqtestGC(results, graph = TRUE)
CHI
```
 
  We now test the independence of the HIV vaccine and placebo group using the chi-square test.  We will state the null and alternative hypotheses for this test.  

$H_0:$ The placebo and aids vaccine treatment is independent of contracting HIV.  

$H_A:$ Having the vaccine is related to contracting the disease.  

The resulting interpretation is the following: if the null hyothesis is indeed true, then the results we just obtained would happen about `r CHI$p.value` percent of the time. 

As a second approach, we will consider Fisher's exact test.  We will state the null and alternative hypotheses for that test.  

$H_0:$ The odds ratio between contracting HIV and having the vaccine contains the number 1.  

$H_A:$ The odds ratio between contracting HIV and having the vaccine does not contain the number 1.

```{r}
FISH <- fisher.test(results)
FISH
```
  The p-value from the Fisher's exact test is `r FISH$p.value` and the confidence interval is `r FISH$conf.int`. Therefore, we must reject the null hypothesis.
  
From these two tests, there is evidence that the results are not independent nor is the odds ratio 1. 
  
## Conclusions

<Summarise your results as to the effectiveness of the vaccine with explanations that could be understood by your mom>
>>>>>>> Inferential
